---
title: "Data Import"
output: github_document
---

```{r, echo = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


I'm an R Markdown document! 

## Data Import: CSVs

Let's import data using the `readr` package.

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


Look at the data

```{r}
litters_df

head(litters_df)
tail(litters_df)
```


Helpful way to view the whole dataset, but don't have it in the rmd file because it creates a new window and creates problems when knit.

```{r,  eval = FALSE}
View(litters_df)
```


The `skimr` package gives a quick summary and overview of the entire dataset. 

```{r}
skimr::skim(litters_df)
```


`read_csv` options..

can treat random numbers as missing observations when data are coded weirdly

Information in the first couple of rows, can skip how many rows at the beginning of the spreadsheet

```{r, eval = FALSE}
read_csv("data/FAS_litters.csv", na = c("", "NA", 999, 88), skip = 2)
```

Update with the col_type argument since it only reads the first 1000 rows and decide what variable type that is, which could be wrong. 


## Other file formats 

We need to read in an excel spreadsheet ...

```{r}
mlb_df = read_excel("data/mlb11.xlsx")
```

```{r, eval = FALSE}
View(mlb_df)
```

```{r}
lotr_words_df = 
  read_excel(
    "data/LotR_Words.xlsx",
    range = "B3:D6"
  )
```

## STill more formats ...

Read in a SAS dataset.

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```

## Data export 

```{r}
write_csv(lotr_words_df, file = "results/lotr_words_df.csv")
```

Use relative paths so that things are well-organized and tightly-bundled into one place

## Why not base r???

```{r}
dont_do_this_df = read.csv("data/FAS_litters.csv")
```

* Import as a dataframe and print the entire thing.
* Categorical or factor instead of character variables; different choice; tibble better than df; read_csv is almost always better than read.csv

